Fitting Synchrotron Models to Gamma-Ray Burst Data from INTEGRAL/SPI

### 2021-12-02 src/ms.tex
fix figure
fix label name
add corner plot

\label{firstpage}
% Abstract with filler text
\abstract {
  In the last years it has been shown, that it is needed to fit physical models to Gamma-Ray Burst (GRB) data to get more insights into their emission mechanism. For the energy range between ~ 10 keV to ~ 10 MeV Fermi/GBM has been the main instrument.... In this paper we show how we can add INTEGRAL/SPI to this analysis. INTEGRAL/SPI covers a similar energy range like Fermi/GBM and has a very good energy resolution of 2.2 keV at 1.3 MeV. Therefore INTEGRAL/SPI is an ideal instrument to precisely measure the curvature of the spectrum. In this work we present PySPI, a new analysis software we developed to analyze GRB data from SPI, and apply it to the GRB 120711A. We show that the GBM and the SPI data of this GRB can be fitted with a physical synchrotron model and that combining both, reduces the allowed parameter space.
}

@@

Since shortly after the start of INTEGRAL, it is known that there are spurious events in the SPI data around 1.5 MeV. According to (cite Roques) these spurious events are photons with small energy (<100 keV) that get detected at a higher energy due to saturation of the ??? electronics by previous high energy deposition. It is also known, that these spurious events do not show up if one looks only at events with a detection also in the Pulse Shape Discriminator (PSD) electronics. The reason for this is that the PSD electronics has an low energy threshold of roughly 450 keV (the exact value has been changed a few times during the mission). Therefore only events that deposit more than this low energy threshold in the Germanium crystal can trigger this electronic, which eliminates the <100 keV events that are detected at the wrong energies by the AFEE electronics. In Fig. \ref{fig:electronic} one can see that the feature at 1.5 MeV is nicely visible in the non-PSD events but is missing in the PSD events. Even though this problem is the most significant in the area around 1.5 MeV it is also important at lower energies >400 keV.

@@

In PySPI every Germanium detector is treated as an independent detector. The workflow during a fit step is the following:
\begin{itemize}
  \item Sample model parameters
  \item Calculate model flux and responses individually for all detectors for the source position
  \item Fold model with responses to get predicted model counts in all detectors
  \item Calculate log-likelihood for all detectors
  \item Sum these log-likelihoods to get the total log-likelihood of SPI
\end{itemize}
A sampling algorithm like emcee (cite) or multinest (cite) is used to iterate this procedure, in order to determine the posterior distributions.\\
This workflow is very similar, like for example for Fermi/GBM, because fundamentally instruments like INTEGRAL/SPI and Fermi/GBM are very similar. They both consist of individual detectors that have different responses for a given source position. In GBM this is due to the different pointing directions of the detectors and in SPI due to the different occultation by the mask. Therefore there is no reason to treat SPI specially, just because it is a coded mask instrument. The information the mask encodes into the data is automatically included by using the different responses for the detectors.

@@

\begin{figure}
  \begin{centering}
    \begin{subfigure}[b]{1\textwidth}
      \begin{centering}
        \includegraphics{figures/spi_lightcurve_ts.pdf}
        \caption{Light curve for SPI detector 0 and GRB120711A.}
        \label{fig:ts_spi}
      \end{centering}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{1\textwidth}
      \begin{centering}
        \includegraphics{figures/gbm_lightcurve_ts.pdf}
        \caption{Light curve for GBM detector n2 and GRB120711A.}
        \label{fig:ts_gbm}
      \end{centering}
    \end{subfigure}
  \end{centering}
\end{figure}

@@

>>>>

\label{firstpage}
% Abstract with filler text
\abstract {
  In the last years it has been shown, that it is needed to fit physical models to Gamma-Ray Burst (GRB) data to get more insights into their emission mechanism. For the energy range between ~ 10 keV to ~ 10 MeV Fermi/GBM has been the main instrument.... In this paper we show how we can add INTEGRAL/SPI to this analysis. INTEGRAL/SPI covers a similar energy range like Fermi/GBM and has a very good energy resolution of 2.2 keV at 1.3 MeV. Therefore INTEGRAL/SPI is an ideal instrument to precisely measure the curvature of the spectrum. In this work we present PySPI, a new analysis software we developed to analyze GRB data from SPI, and apply it to the GRB 120711A. We show that PySPI improves the analysis of SPI data, for GRBs, compared to the official analysis software. Also we show that the GBM and the SPI data of this GRB can be fitted well with a physical synchrotron model and that combining both, reduces the allowed parameter space.
}

@@

Since shortly after the start of INTEGRAL, it is known that there are spurious events in the SPI data around 1.5 MeV. According to (cite Roques) these spurious events are photons with small energy (<100 keV) that get detected at a higher energy due to saturation of the ??? electronics by previous high energy deposition. It is also known, that these spurious events do not show up if one looks only at events with a detection also in the Pulse Shape Discriminator (PSD) electronics. The reason for this is that the PSD electronics has an low energy threshold of roughly 450 keV (the exact value has been changed a few times during the mission). Therefore only events that deposit more than this low energy threshold in the Germanium crystal can trigger this electronic, which eliminates the <100 keV events that are detected at the wrong energies by the AFEE electronics. In Fig. \ref{fig:electronic} one can see that the feature at 1.5 MeV is nicely visible in the non-PSD events but is missing in the PSD events. Even though this problem is the most significant in the area around 1.5 MeV it is also important at lower energies >400 keV. In PySPI one can select the energy range in which only the PSD events should be used, to avoid including spurious events. To account for the larger dead time of the PSD electronics, an effective area correction can be either fix to 85 \% (cite) or treated as a free parameter in the fit.

@@

In PySPI every Germanium detector is treated as an independent detector. The workflow during a fit step is the following:
\begin{itemize}
  \item Sample model parameters
  \item Calculate model flux and responses individually for all detectors for the source position
  \item Fold model with responses to get predicted model counts in all detectors
  \item Calculate log-likelihood for all detectors
  \item Sum these log-likelihoods to get the total log-likelihood of SPI
\end{itemize}
A sampling algorithm like emcee (cite) or MultiNest (cite) is used to iterate this procedure, in order to determine the posterior distributions.\\
This workflow is very similar, like for example for GBM, because fundamentally instruments like SPI and GBM are very similar. They both consist of individual detectors that have different responses for a given source position. In GBM this is due to the different pointing directions of the detectors and in SPI due to the different occultation by the mask. Therefore there is no reason to treat SPI specially, just because it is a coded mask instrument. The information the mask encodes into the data is automatically included by using the different responses for the detectors.

@@

\begin{figure}
  \begin{centering}
    \includegraphics[width=1\linewidth]{figures/spi_lightcurve_ts.pdf}
    \includegraphics[width=1\linewidth]{figures/gbm_lightcurve_ts.pdf}
    \label{fig:time_selection}
    \caption{Light curve for GRB120711A.}
  \end{centering}
\end{figure}

@@

\begin{figure}
  \begin{centering}
    \includegraphics[width=1\linewidth]{figures/pyspi_vs_gbm_band.pdf}
    \label{fig:corner_gb_pyspi_band}
    \caption{Corner plot for the SPI fit with PySPI and the GBM fit. In this fit we used a Band function as spectral model.}
  \end{centering}
\end{figure}
